### 2021-08-01 paper/paper.tex
first draft of summary

\trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
high-order numerical simulations of hyperbolic partial differential equations (PDEs).
in Julia \cite{bezanson2017julia}.

@@

\todo{TODO %TODO Summary
\begin{itemize}
  \item Julia is suited for simulation-focused scientific computing but some
        things need to be improved/demonstrated for HPC
  \item Julia allows us to do some cool stuff in \trixi
  \item Learning a new programming language has some overhead but it pays off
\end{itemize}}

@@

\todo{ % TODO
\begin{itemize}
  \item Full paper: 5--10 pages
  \item Prepare repro repo including scripts for \trixi and FLUXO
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks
  \item Shall we re-create the schematic overview in tikz?
  \item Shall we compare \trixi (in terms of structure, not performance!)
        to other open source frameworks such as MFEM,
        NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...? Definitely not
        explicitly to not make people angry, but maybe implicitly without
        mentioning their names?
        \begin{itemize}
          \item Everything in Julia, not a two language problem!
          \item Other things? Gregor will think about it!
        \end{itemize}
  \item Are there other aspects we should discuss?
  \item Check affiliations
  \item Check order of authors
  \item Maybe include some cool pictures? Give a flavor of the kind of simulations
        you can do with \trixi? Pictures trigger people! For example after the
        list of main features?
  \item Summary of community, collaborators, and users: GitHub statistics, contributors,
        Vikram and Chandrashekar, Steven Frankel and Philipp Öffner (maybe), ...
        Related to the relatively large number of students!
        $\to$ Jesse will make a PR!
  \item A simple example where we really use multiple dispatch?
        \begin{itemize}
          \item Just pass functions around
          \item Dispatch on parameters of DG struct
        \end{itemize}
\end{itemize}
\begin{quote}
  Compared to an extended abstract, a full paper presents more of the background
  and context motivating the work. It compares the work to other approaches taken
  in the field and gives some additional insights on the conference contribution.
  Use cases back up the work by showing how it can be used.
\end{quote}}

>>>>

\trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
high-order numerical simulations of hyperbolic partial differential equations (PDEs)
in Julia \cite{bezanson2017julia}.

@@

We have presented \trixi, a Julia library for adaptive high-order numerical
simulations of hyperbolic PDEs. As researchers in numerical analysis and
scientific computing, our goals were to create a framework that is extendable,
easy to use, and fast (in this particular order). Making use of Julia's strengths,
we have been successful based on recent publications making use of \trixi,
the number of students and researchers working with \trixi, and serial performance
comparisons with established Fortran codes. Having worked on \trixi for a bit
more than one year allows us to give an assessment of Julia for simulation-focused
scientific computing.

Based on our experience, we consider Julia to be suitable for simulation-focused
scientific computing, in particular for hyperbolic PDEs, computational fluid
dynamics, and related problems --- at least on the scale of shared memory
parallelism. The scalability of high-order methods for hyperbolic PDEs written
in Julia to high performance computing applications still needs to be demonstrated.
However, we do not expect unsolvable problems there since Julia is not generically
slower than traditional compiled HPC programming languages. Nevertheless, it
appears to be more complicated to scale Julia to HPC while still keeping the
code as simple, clean, and flexible as it is now.

Learning a new programming language requires of course some effort and needs some
time. Nevertheless, it pays off in our experience and we do not want to miss the
nice features of \trixi enabled by Julia and its ecosystem. Thus, while there is
of course no need to switch to Julia if people are satisfied with their existing
tools, we encourage researchers to try out Julia and stay for a while. The Julia
community (on the Julia Discourse forum\footnote{\url{https://discourse.julialang.org/}}
or Slack/Zulip workspaces) is usually welcoming and helpful, both for new and
experienced users. Making use of existing learning material and asking questions
appears to be important since a new programming style is necessary to make full
use of all of Julia's strengths.

@@

\todo{ % TODO
\begin{itemize}
  \item Full paper: 5--10 pages
  \item Prepare repro repo including scripts for \trixi and FLUXO
  \item Restructure non-conservative stuff and add MHD benchmarks
  \item Shall we re-create the schematic overview in tikz?
  \item Shall we compare \trixi (in terms of structure, not performance!)
        to other open source frameworks such as MFEM,
        NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...? Definitely not
        explicitly to not make people angry, but maybe implicitly without
        mentioning their names?
        \begin{itemize}
          \item Gregor will think about it!
        \end{itemize}
  \item Are there other aspects we should discuss?
  \item Check affiliations
  \item Check order of authors
  \item Maybe include some cool pictures? Give a flavor of the kind of simulations
        you can do with \trixi? Pictures trigger people! For example after the
        list of main features?
  \item Summary of community, collaborators, and users: GitHub statistics, contributors,
        Vikram and Chandrashekar, Steven Frankel and Philipp Öffner (maybe), ...
        Related to the relatively large number of students!
        $\to$ Jesse will make a PR!
  \item A simple example where we really use multiple dispatch?
        \begin{itemize}
          \item Just pass functions around
          \item Dispatch on parameters of DG struct
        \end{itemize}
\end{itemize}
\begin{quote}
  Compared to an extended abstract, a full paper presents more of the background
  and context motivating the work. It compares the work to other approaches taken
  in the field and gives some additional insights on the conference contribution.
  Use cases back up the work by showing how it can be used.
\end{quote}}
